---
title: "table_construction_test"
author: "Mike Stefanuk"
date: "August 6, 2018"
output: word_document
---

Lost my old document on creating tables to feed to random forest. Here is v2.

Stage 1: Create some test data...

```{r test data, echo = FALSE}
# test data
avhrrp <- raster::brick(array(rnorm(n = (5*5*12), mean = 1.0, sd = 0.2), c(5,5,12)))
avhrrp <- raster::brick(raster::stack(avhrrp, avhrrp, avhrrp, avhrrp, avhrrp))

temp <- raster::brick(array(rnorm(n = (5*5*12), mean = 15, sd = 5), c(5,5,12)))
temp <- raster::brick(raster::stack(temp, temp, temp, temp, temp))

prcp <- raster::brick(array(rnorm(n = (5*5*12), mean = 2, sd = 10), c(5,5,12)))
prcp <- raster::brick(raster::stack(prcp, prcp, prcp, prcp, prcp))
prcp[prcp < 0] <- 0

indx <- raster::brick(array(rnorm(n = (5*5*12), mean = 50, sd = 15), c(5,5,12)))
indx <- raster::brick(raster::stack(indx, indx, indx, indx, indx))

# set names of test data to months and years 
months <- zoo::as.yearmon(seq(from = as.Date("1901-01-15"),
                              to = as.Date("1905-12-15"),
                              by = "month"))

avhrrp <- raster::setZ(avhrrp, months, name = "time")
names(avhrrp) <- months

temp <- raster::setZ(temp, months, name = "time")
names(temp) <- months

prcp <- raster::setZ(prcp, months, name = "time")
names(prcp) <- months

indx <- raster::setZ(indx, months, name = "time")
names(indx) <- months

rm(months)

```

Ok, test data is created, names are set and z values are set.

Stage 2: play with table creation

```{r table creation 1, echo = FALSE, message = FALSE}
# convert avhrrp to an ordered column
a1 <- raster::as.data.frame(avhrrp)
a2s <- stack(a1) # results are ok but formatting is annoying
a2u <- unlist(a1) # produces a named vector but order of months is wrong
a2r <- reshape2::melt(a1, na.rm = FALSE) # works great!!

a <- reshape2::melt(raster::as.data.frame(avhrrp), na.rm = FALSE)
b <- reshape2::melt(raster::as.data.frame(temp),   na.rm = FALSE)
c <- reshape2::melt(raster::as.data.frame(prcp),   na.rm = FALSE)
d <- reshape2::melt(raster::as.data.frame(indx),   na.rm = FALSE)

# create dataframe
dframe <- data.frame(time = a$variable, avhrrp = a$value,
                     temp = b$value, prcp = c$value, indx = d$value)

# cleanup
rm(a1,a2s, a2u, a2r, a, b, c, d)

```

This method seems to make a nice logical dataframe.

Can I make a function to construct such a dataframe with different numbers of inputs?

```{r function test, echo = FALSE}
# make a function to create a dataframe by taking different numbers of arguments
maker <- function(in.data = NULL, na.rm = FALSE){
  
  # check that the inputs are a list
  if(class(in.data) != "list"){
    stop("input data must be given as a list")
  }
  
  # create an output list
  out <- list()
  
  # loop through items on list
  for(i in 1:length(names(in.data))){
    # convert raster brick to dataframe
    o1 <- raster::as.data.frame(in.data[[i]])
    
    # add pixel number identifier to dataframe
    o2 <- data.frame(pix = rownames(o1), o1)
    
    # convert dataframe to long format
    o3 <- reshape2::melt(o2, id = "pix")
    
    # write value strings
    out[i] <- list(o3$value)
    
  }
  
  # convert to dataframe
  out <- as.data.frame(out)
  
  # assign rownames to dataframe
  ## month.year.pixel
  rownames(out) <- paste0(o3[,2], ".", o3[,1])
  
  if(na.rm == TRUE){
    # filter out rows where the first value column is NA
    out <- dplyr::filter(out, !is.na(names(out)[1]))
  }
  
  # assign column names
  colnames(out) <- names(in.data)
  
  # return result
  return(out)
}

```

```{r function testing}

# create a list of inputs, avhrrp being first
l <- list(avhrr = avhrrp, tmp = temp, prc = prcp, ind = indx)

# run function and view header
t1 <- maker(l)

# view header
head(t1)

```

Now, let's see if this can be made to work with randomForest

Requirements

- y = response = avhrrp (might be easiest to have this in its own df)

- x = predictor = climate (these should be combined into one df)

- making a single dataframe as above is likely not the best way to go

Stage 3 : Random Forest Functionality

```{r randomForest, echo = FALSE}

t2 <- randomForest::tuneRF(x = data.frame(t1$tmp, t1$prc, t1$ind), 
                           y = t1$avhrr)

t3 <- randomForest::randomForest(data = t1, avhrr~.,
                                 #do.trace = (nrow(t1)/100), # prints some tree info
                                 importance = TRUE, 
                                 mtry = 2, 
                                 replace = TRUE)

randomForest::importance(t3)

randomForest::partialPlot(t3, pred.data = t1, x.var = "tmp")
randomForest::partialPlot(t3, pred.data = t1, x.var = "prc")
randomForest::partialPlot(t3, pred.data = t1, x.var = "ind")

```

